Response to reviewers

### 2016-11-15 docs/AIAA 2016 highway platooning/response1/response1.tex
minor changes to revision
updated talk
updated revision

Comments and questions from reviewers have been either bolded or italicized. The bolded comments and questions have been directly addressed below. All comments have been addressed in the paper as much as possible without major changes to the core of the paper.

@@

However, any forecast, however uncertain it may be, can be incorporated into the cost map. Given the cost map, air highways can then be placed. The exact process of converting everything including population density, geography, forecast information, etc. to a cost map is not discussed in the paper. One primary purpose of the paper is to highlight that air highway placement can be done in real-time given the cost map.

@@

This is great point. The cost map would be changed based on many factors such as weather forecasts (eg. impending hurricane), commercial air traffic, etc. This could either be done manually or automatically. In terms of communicating this information, all vehicles and users could have an interface showing the current highway map.

@@

Under the assumption that UAVs can only fly when they're guaranteed to be safe, then as the demand for flying increases, vehicles would no longer be guaranteed-safe when they are free. The platooning structure allows safety guarantees to be made, and the multi-UAV system to be easily monitored, thereby allowing more vehicles to fly at once. Whether the vehicles are allowed to be free during their entire flight would be subject to traffic rules to be established by government agencies (or some centralized governing body).

@@

Definitely, being part of the highway structure has a cost to the individual vehicle. However, remaining free may not be an option depending on traffic rules. When the number of vehicles increases to a point where all vehicles being free is no longer manageable, each vehicle would need to use the highways, and effectively ``pay'' for a structure that guarantees safety.

@@

Strictly speaking there is no maximum. However, practically speaking it is unclear what the best size for a platoon would be on a single highway. Further research would be needed to understand this.

\item\textit{No future work in Conclusions.  The Conclusions section is only for conclusions that can be drawn directly from the manuscript.}
\end{itemize}

\textbf{Additional Minor Comments:}
\begin{enumerate}
\item \textit{Please ensure that the captions of all figures are not longer than two lines.  Additional text can be included in the main body of the pager.}
\item \textit{Please re-define all acronyms in the Conclusions.}
\item \textit{The Conclusions section is reserved for conclusions that can be drawn from the body of the paper.  Please delete the last sentence since it is referring to future work.}
\item \textit{DOI numbers are required for the references whenever available}
\item \textit{AIAA conference papers need paper numbers}
\end{enumerate}

@@

\textit{A well-written paper on a topic of increasing importance.  The issue of traffic flow management as applied to low-altitude, small UAS is undervalued by many in ATM and aviation.  It is encouraging to see solid work being done in this area.}

\textit{Overall the presentation of the paper is clear.  However, as a reader, I was left with several (small) open questions.  Those questions need not be deeply answered by the paper, however, I think addressing them explicitly (even if to state they will not be detailed) will provide a more complete feeling to the paper.}

\begin{enumerate}
\item\textit{This work is like platooning research from traditional TFM.  The problem with those solutions was typically that through the implementation of platooning, often there was such a loss of efficiency to form the platoon, that it was ultimately not worth while.  The authors should probably look at some of that previous work for citation rather than relying solely on ground vehicle platooning.}

@@

This point is very important and requires further investigation. As long as multiple conflicts do not occur, pairwise collision avoidance would be sufficient to keep vehicles safe in an unstructured environment. However, it is difficult to predict whether multiple conflicts would occur, and in addition, unstructured flight can be very unintuitive to monitor.

@@

One assumption implicitly made in this paper is that the mechanism for exiting the altitude range can be abstracted or summarized into just a time-to-exit-altitude-range parameter.

\item\textit{A few sentences about the policy implications may help ground the reader in terms of how such a system might find its way in to the airspace in the future.  Is this platooning a requirement of operators?  Optional?  If optional, what is the critical mass that makes it effective?}

\item\textit{Should your state diagram (Figure 4) have a terminal state of some kind?  Like a "Closed" state where they just aren't flying anymore?  Perhaps reachable from "Free" and "Faulty"?  Otherwise it seems like all vehicles are in the air at all times.  If so, you may want to consider an entry point to the state diagram as well.  Where can the vehicles start out?}
\end{enumerate}

\textit{Again, I don't think all of these questions need to be answered in detail, but addressing them will strengthen the paper.}

\textit{The only style point I have, is in that the introduction sounds a bit too "salesman" like.  In the first paragraph, terms like "important" and "human imagination" feel like they may be a tad overplayed.  Later terms like "cannot be denied" and perhaps a couple others detract from the technical work in some way.  Not required that these be changed, just a style note from one reader.}

@@

\textit{The paper draws inspiration from the Automated Highway Systems (AHS) concept developed at Berkeley in the 90s. The paper proposes the introduction of structure, namely Air Highways and platoons, to airspace control, which is bound to be congested in a near future because of the drone revolution. This seems to be the contribution of the paper.}

\textit{The problem is that there is no technical justification (or discussion) for this approach. True, that corridors (or Air Highways) make sense, especially in congested areas. But platooning, as proposed, is difficult to justify (in fact there is no justification for platoons in the paper).}

@@

@@

\item \textit{“liveness” – Please cite optimal control work on liveness.}

@@

\item \textit{''HJ reachability computations are intractable for a large number of vehicles'' – Actually it is not tractable for one single vehicle if we consider dynamics evolving in a high dimension state-space.}

@@

This paper is not meant to provide an exhaustive solution. It illustrates that having a highway structure in the airspace allows HJ reachability, a method that normally can only analyze small-scale systems, to be used for large-scale systems. There are definitely those issues mentioned in the comment, as well as many other challenges that need to be overcome. Hopefully this paper can provide a starting point on which future research can be built, and make future research problems more concrete.
\end{itemize}

@@

The paper does not explicitly address the problem of finding common pathways, but the air highway placement method presented here naturally leads to common pathways as seen in figures 2 and 3

@@

II.A The air highway placement problem
\item \textit{``... to each air highway we assign a speed of travel'' – shouldn’t this be a control setting? This varies a lot with the type of UAVs.}
\item \textit{``...air highway locations should ideally be able to be recomputed in real-time when necessary in order to update airspace constraints on-the fly'' – how practical is this under the limitations discussed above?}
\item \textbf{ ``... air highway placement problem, and'' – the title can be misleading. The problem is not about finding all highways for a given region. In fact the formulation considers just one “source” and multiple destinations. It is unclear how it would work for the problem of multiple ``sources'' and destinations}

@@

\item \textit{Cost function – shouldn’t the cost function depend on the type of UAV?}
\end{itemize}

II.B. The Eikonal equation
\begin{itemize}
\item \textit{``sequence of paths'' – same problem as above for highways}
\item \textbf{``It is well known that the solution to the Eikonal equation (5)'' – this statement is misleading. This is only true if the dynamics of the UAV are of a special form. This is poorly explanined in this paragraph. It neglects to discuss how the Eikonal equation is derived for a special form of dynamics. Moreover, in the tradition of optimal control theory V is called the value function.}

@@

\item \textit{``Given a single origin point po, the optimal cumulative cost function V (pd) can be computed. Suppose M different destination points. Then, M different optimal paths Pi; i = 1; : : : ;M are obtained from V'' – True, but misleading. In fact any number of M will do since the value function is defined for the whole state space.}
\item \textit{ Equation (3) – Are there any assumptions on the cost function? If not it may happen that V may not be differentiable. What about the interpretation of equation (5) for this case?}
\end{itemize}

E. Real-Time Highway Location Updates
\begin{itemize}
\item \textit{``Since (5) can be solved in approximately 1 second, the air highway placement process can be redone in real-time if the cost'' – True, but the results will no longer satisfy the Principle of Optimality which is at the root of the derivation of the Eikonal equation.}
\item \textit{On the other hand, depending on for instance the time of day, it may be most desirable to fly in different regions of the airspace'' – Again, if the cost varies with time, then the Eikonal equation no longer solves for the value function V.}
\item \textbf{``We add to the collection of waypoints the first point on the path at which the heading changes by some threshold C, and repeat this process along the entire path.'' – The paths arising from the Eikonal equation can have kinks and jumps in curvature. How to address that?}

As mentioned, post-processing is not the focus of this paper.
\end{itemize}

@@

\item \textit{``... problem of collision avoidance among N vehicles cannot be tractably solved using traditional dynamic programming approaches because the computation complexity of these approaches scales exponentially with the number ...'' – This needs clarification. What are the information structures for this problem?}

\item \textit{``liveness and safety of the vehicles in a tractable manner'' – liveness and safety are not properly defined.}
\item \textit{''We consider a platoon of vehicles to be a group of $M \le N$ vehicles, ...'' – This is not clear.}
\item \textit{''If the vehicle is part of platoon, then it checks safety...” – what it the vehicle is the leader of a platoon? This case is addressed in Figure 5, but not here.}
\item \textit{``... have a liveness component, which specifies a set of states that the vehicle aims to reach, and a safety component, which specifies a set of states that it must avoid.'' – This is not a definition of liveness.}

B. Hamilton-Jacobi reachability
\item \textit{Equation (10) – Some of the symbols in the equation were not defined.}
\item \textbf{What is the difference between the HJB and the HJ equations?}
HJ generally refers to either HJB or HJI. The HJB equation is used when the dynamics only involves a control, and the HJI equation is used when the dynamics involve both a control and a disturbance.
\item \textit{Figure 6 – What is the zero plane and where is it in the picture?}
\end{itemize}

C. Reachability based controllers
\begin{itemize}
\item \textit{``Reachability analysis is useful for constructing controllers ...'' – This does not say much. Please elaborate on it.}
\item \textit{``The system dynamics with which the reachable set is computed provide additional flexibility when using reachability to construct controller...'' – Not clear.}
\item \textbf{Aren’t all calculations made much simple because you are in an air highway? Please elaborate on this.}

@@

\item \textit{``liveness controller described in the previous sections when it is not in any danger of collision with other vehicles'' – As before, the concept of liveness has not been introduced before, so it may be difficult to the reader to understand what is going on.}
\item \textit{``3. Avoiding collisions'' – What happens when a potential collision is detected and there is no solution to the obstacle avoidance problem? Not clear from the text if the approach will work for the case when multiple UAVs are using safety controllers? This is even more interesting for the case of UAVs in a platoon. It is not clear from the text if the safety set is defined only for motions along the highway, or if there are additional degrees of freedom.}
\end{itemize}

D. Other controllers
\begin{itemize}
\item \textit{``$\hat d$ is the highway’s direction of travel.'' – The direction of travel is not properly defined.}
\end{itemize}

E. Safety analysis
\begin{itemize}
\item \textit{The discussion seems to overlook the definition of the information structures for the problem. The same happens with considerations about sensor performance.}
\item \textit{``Given that vehicles within a platoon are safe with respect to each other, each platoon can be treated as a single vehicle, and perform collision avoidance with other platoons. By treating each platoon as a single unit, we can reduce the number of individual vehicles that need to check for safety against each other, reducing overall computation burden.'' – This seems to be in contradiction with the example with an intruder vehicle.}
\end{itemize}

F. Numerical simulations
\begin{itemize}
\item \textit{What are the information structures for these problems?}
\item \textit{The maneuvers described in Fig 9 seem to go against our intuition. A more detailed description of what is going on is needed.}
\end{itemize}

IV. Conclusions
\begin{itemize}
\item \textit{``Many complex maneuvers can be performed using only a few reachability sets.'' – This phrase does not say much.}
\end{itemize}

>>>>

@@

However, any forecast, however uncertain it may be, can be incorporated into the cost map. Given the cost map, air highways can then be placed. The exact process of converting everything including population density, geography, forecast information, etc. to a cost map is not discussed in the paper. One primary purpose of the paper is to highlight that air highway placement can be done in real-time given the cost map. The text in the paper has been updated to address this comment.

@@

This is great point. The cost map would be changed based on many factors such as weather forecasts (eg. impending hurricane), commercial air traffic, etc. This could either be done manually or automatically. In terms of communicating this information, all vehicles and users could have an interface showing the current highway map. The text in the paper has been updated to address this comment.

@@

Under the assumption that UAVs can only fly when they're guaranteed to be safe, then as the demand for flying increases, vehicles would no longer be guaranteed-safe when they are free. The platooning structure allows safety guarantees to be made, and the multi-UAV system to be easily monitored, thereby allowing more vehicles to fly at once. Whether the vehicles are allowed to be free during their entire flight would be subject to traffic rules to be established by government agencies (or some centralized governing body). The text in the paper has been updated to address this comment.

@@

Definitely, being part of the highway structure has a cost to the individual vehicle. However, remaining free may not be an option depending on traffic rules. When the number of vehicles increases to a point where all vehicles being free is no longer manageable, each vehicle would need to use the highways, and effectively ``pay'' for a structure that guarantees safety. The text in the paper has been updated to address this comment.

@@

Strictly speaking there is no maximum. However, practically speaking it is unclear what the best size for a platoon would be on a single highway. Further research would be needed to understand this. The text in the paper has been updated to address this comment.

\item\textbf{No future work in Conclusions.  The Conclusions section is only for conclusions that can be drawn directly from the manuscript.}

The text in the paper has been updated to address this comment.
\end{itemize}

\textbf{Additional Minor Comments:}
\begin{enumerate}
\item \textbf{Please ensure that the captions of all figures are not longer than two lines.  Additional text can be included in the main body of the pager.}
\item \textbf{Please re-define all acronyms in the Conclusions.}
\item \textbf{The Conclusions section is reserved for conclusions that can be drawn from the body of the paper.  Please delete the last sentence since it is referring to future work.}
\item \textbf{DOI numbers are required for the references whenever available}
\item \textbf{AIAA conference papers need paper numbers}
\end{enumerate}
The text in the paper has been updated to address these comments.

@@

\textbf{A well-written paper on a topic of increasing importance.  The issue of traffic flow management as applied to low-altitude, small UAS is undervalued by many in ATM and aviation.  It is encouraging to see solid work being done in this area.}

\textbf{Overall the presentation of the paper is clear.  However, as a reader, I was left with several (small) open questions.  Those questions need not be deeply answered by the paper, however, I think addressing them explicitly (even if to state they will not be detailed) will provide a more complete feeling to the paper.}

\begin{enumerate}
\item\textbf{This work is like platooning research from traditional TFM.  The problem with those solutions was typically that through the implementation of platooning, often there was such a loss of efficiency to form the platoon, that it was ultimately not worth while.  The authors should probably look at some of that previous work for citation rather than relying solely on ground vehicle platooning.}

Platooning and flying on air highways in general can be seen as a cost that the system has to ``pay'' to ensure tractable analysis. This point has been added to the paper.

@@

This point is very important and requires further investigation. As long as multiple conflicts do not occur, pairwise collision avoidance would be sufficient to keep vehicles safe in an unstructured environment. However, it is difficult to predict whether multiple conflicts would occur, and in addition, unstructured flight can be very unintuitive to monitor. This point has been emphasized slightly in the paper.

@@

One assumption implicitly made in this paper is that the mechanism for exiting the altitude range can be abstracted or summarized into just a time-to-exit-altitude-range parameter. This point has been clarified in the paper.

\item\textbf{A few sentences about the policy implications may help ground the reader in terms of how such a system might find its way in to the airspace in the future.  Is this platooning a requirement of operators?  Optional?  If optional, what is the critical mass that makes it effective?}

This comment is in general similar to the first comment, and this point has been added to the paper.

\item\textbf{Should your state diagram (Figure 4) have a terminal state of some kind?  Like a "Closed" state where they just aren't flying anymore?  Perhaps reachable from "Free" and "Faulty"?  Otherwise it seems like all vehicles are in the air at all times.  If so, you may want to consider an entry point to the state diagram as well.  Where can the vehicles start out?}

Modified figure caption to say that vehicles begin in the ``Free'' mode initially.
\end{enumerate}

\textbf{Again, I don't think all of these questions need to be answered in detail, but addressing them will strengthen the paper.}

\textbf{The only style point I have, is in that the introduction sounds a bit too "salesman" like.  In the first paragraph, terms like "important" and "human imagination" feel like they may be a tad overplayed.  Later terms like "cannot be denied" and perhaps a couple others detract from the technical work in some way.  Not required that these be changed, just a style note from one reader.}

Agreed. Text has been changed slightly to address this.

@@

\textbf{The paper draws inspiration from the Automated Highway Systems (AHS) concept developed at Berkeley in the 90s. The paper proposes the introduction of structure, namely Air Highways and platoons, to airspace control, which is bound to be congested in a near future because of the drone revolution. This seems to be the contribution of the paper.}

\textbf{The problem is that there is no technical justification (or discussion) for this approach. True, that corridors (or Air Highways) make sense, especially in congested areas. But platooning, as proposed, is difficult to justify (in fact there is no justification for platoons in the paper).}

@@

The text in the paper has been modified to address all of these major comments.

@@

\item \textbf{“liveness” – Please cite optimal control work on liveness.}

@@

\item \textbf{''HJ reachability computations are intractable for a large number of vehicles'' – Actually it is not tractable for one single vehicle if we consider dynamics evolving in a high dimension state-space.}

@@

This paper is not meant to provide an exhaustive solution. It illustrates that having a highway structure in the airspace allows HJ reachability, a method that normally can only analyze small-scale systems, to be used for large-scale systems. There are definitely those issues mentioned in the comment, as well as many other challenges that need to be overcome. Hopefully this paper can provide a starting point on which future research can be built, and make future research problems more concrete. This has been better emphasized in the revised paper.
\end{itemize}

@@

The paper does not explicitly address the problem of finding common pathways, but the air highway placement method presented here naturally leads to common pathways as seen in figures 2 and 3. This point has already been stated in the paper

@@

II.A The air highway placement problem
\item \textbf{``... to each air highway we assign a speed of travel'' – shouldn’t this be a control setting? This varies a lot with the type of UAVs.}

The text has been updated to emphasize that details like this are not the focus of the paper.

\item \textbf{``...air highway locations should ideally be able to be recomputed in real-time when necessary in order to update airspace constraints on-the fly'' – how practical is this under the limitations discussed above?}

The changes to the cost map should not be continuously varying in time. This has been clarified in the revision.

\item \textbf{ ``... air highway placement problem, and'' – the title can be misleading. The problem is not about finding all highways for a given region. In fact the formulation considers just one “source” and multiple destinations. It is unclear how it would work for the problem of multiple ``sources'' and destinations}

@@

\item \textbf{Cost function – shouldn’t the cost function depend on the type of UAV?}

Perhaps. In this case, perhaps a different set of highways should be created, or somehow different cost maps should be merged together. However this point is out of the scope of the paper.

\end{itemize}

II.B. The Eikonal equation
\begin{itemize}
\item \textbf{``sequence of paths'' – same problem as above for highways}
\item \textbf{``It is well known that the solution to the Eikonal equation (5)'' – this statement is misleading. This is only true if the dynamics of the UAV are of a special form. This is poorly explanined in this paragraph. It neglects to discuss how the Eikonal equation is derived for a special form of dynamics. Moreover, in the tradition of optimal control theory V is called the value function.}

@@

\item \textbf{``Given a single origin point po, the optimal cumulative cost function V (pd) can be computed. Suppose M different destination points. Then, M different optimal paths Pi; i = 1; : : : ;M are obtained from V'' – True, but misleading. In fact any number of M will do since the value function is defined for the whole state space.}
\item \textbf{ Equation (3) – Are there any assumptions on the cost function? If not it may happen that V may not be differentiable. What about the interpretation of equation (5) for this case?}

Good point regarding differentiability. The solutions of PDEs in this paper are all viscosity solutions. The text has been changed accordingly with a citation on viscosity solutions added.
\end{itemize}

E. Real-Time Highway Location Updates
\begin{itemize}
\item \textbf{``Since (5) can be solved in approximately 1 second, the air highway placement process can be redone in real-time if the cost'' – True, but the results will no longer satisfy the Principle of Optimality which is at the root of the derivation of the Eikonal equation.}

\item \textbf{On the other hand, depending on for instance the time of day, it may be most desirable to fly in different regions of the airspace'' – Again, if the cost varies with time, then the Eikonal equation no longer solves for the value function V.}
\item \textbf{``We add to the collection of waypoints the first point on the path at which the heading changes by some threshold C, and repeat this process along the entire path.'' – The paths arising from the Eikonal equation can have kinks and jumps in curvature. How to address that?}

As mentioned, post-processing is not the focus of this paper. The text has been changed to emphasize this.
\end{itemize}

@@

\item \textbf{``... problem of collision avoidance among N vehicles cannot be tractably solved using traditional dynamic programming approaches because the computation complexity of these approaches scales exponentially with the number ...'' – This needs clarification. What are the information structures for this problem?}

\item \textbf{``liveness and safety of the vehicles in a tractable manner'' – liveness and safety are not properly defined.}
\item \textbf{''We consider a platoon of vehicles to be a group of $M \le N$ vehicles, ...'' – This is not clear.}
\item \textbf{''If the vehicle is part of platoon, then it checks safety...” – what it the vehicle is the leader of a platoon? This case is addressed in Figure 5, but not here.}
\item \textbf{``... have a liveness component, which specifies a set of states that the vehicle aims to reach, and a safety component, which specifies a set of states that it must avoid.'' – This is not a definition of liveness.}

B. Hamilton-Jacobi reachability
\item \textbf{Equation (10) – Some of the symbols in the equation were not defined.}
\item \textbf{What is the difference between the HJB and the HJ equations?}
HJ generally refers to either HJB or HJI. The HJB equation is used when the dynamics only involves a control, and the HJI equation is used when the dynamics involve both a control and a disturbance. To avoid confusion, all HJI and HJB PDEs are both now referred to as HJ PDEs.

\item \textbf{Figure 6 – What is the zero plane and where is it in the picture?}
\end{itemize}

C. Reachability based controllers
\begin{itemize}
\item \textbf{``Reachability analysis is useful for constructing controllers ...'' – This does not say much. Please elaborate on it.}
\item \textbf{``The system dynamics with which the reachable set is computed provide additional flexibility when using reachability to construct controller...'' – Not clear.}

Details are in later subsections.

\item \textbf{Aren’t all calculations made much simple because you are in an air highway? Please elaborate on this.}

@@

\item \textbf{``liveness controller described in the previous sections when it is not in any danger of collision with other vehicles'' – As before, the concept of liveness has not been introduced before, so it may be difficult to the reader to understand what is going on.}
\item \textbf{``3. Avoiding collisions'' – What happens when a potential collision is detected and there is no solution to the obstacle avoidance problem? Not clear from the text if the approach will work for the case when multiple UAVs are using safety controllers? This is even more interesting for the case of UAVs in a platoon. It is not clear from the text if the safety set is defined only for motions along the highway, or if there are additional degrees of freedom.}
\end{itemize}

D. Other controllers
\begin{itemize}
\item \textbf{``$\hat d$ is the highway’s direction of travel.'' – The direction of travel is not properly defined.}
\end{itemize}

E. Safety analysis
\begin{itemize}
\item \textbf{The discussion seems to overlook the definition of the information structures for the problem. The same happens with considerations about sensor performance.}
\item \textbf{``Given that vehicles within a platoon are safe with respect to each other, each platoon can be treated as a single vehicle, and perform collision avoidance with other platoons. By treating each platoon as a single unit, we can reduce the number of individual vehicles that need to check for safety against each other, reducing overall computation burden.'' – This seems to be in contradiction with the example with an intruder vehicle.}

When there is only one platoon, it made more sense to not treat it as a single unit in order to show a more interesting behavior. The wording of the paper has been changed to say that platoons \textit{can be} treated as single unit \textit{if needed}.
\end{itemize}

F. Numerical simulations
\begin{itemize}
\item \textbf{What are the information structures for these problems?}
\item \textbf{The maneuvers described in Fig 9 seem to go against our intuition. A more detailed description of what is going on is needed.}
\end{itemize}

IV. Conclusions
\begin{itemize}
\item \textbf{``Many complex maneuvers can be performed using only a few reachability sets.'' – This phrase does not say much.}
\end{itemize}
